Distribution Theory and Applications --- Example Sheet 1

### 2020-10-29 distribution_theory/distribution_ES1.tex
why is distribution theory so *****ing DIFFICULT

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\ord}{ord}

@@

\begin{solution}
	We follow the hint: let $\eps > 0$, then we first compute
	\[
	\int_{\norm{x} > \eps} \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \frac{\pdv[2]{x_i} \phi}{\norm{x}^{n-2}}\dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \phi \cdot \pdv[2]{x_i} \norm{x}^{2-n} \dd{x}. 
	\]
	Now, it is easily computed that $\pdv{x_i} \norm{x} = \frac{x_i}{\norm{x}}$, and therefore 
	\begin{align*}
	\pdv[2]{x_i} \norm{x}^{2-n} &= \pdv{x_i} (2-n) x_i \norm{x}^{-n} = (2-n) \qty(\norm{x}^{-n} - n x_i^2 \norm{x}^{-n-2}) \\
	&= (n-2)\norm{x}^{-n} \qty(n \qty(\frac{x_i}{\norm{x}})^2 - 1). 
	\end{align*}
\end{solution}

\end{document}

>>>>

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\prv}{p.\!v.}

@@

\begin{solution}
	We follow the hint: let $\eps > 0$, then we first compute
	\[
	\int_{\norm{x} > \eps} \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \frac{\pdv[2]{x_i} \phi}{\norm{x}^{n-2}}\dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \phi \cdot \pdv[2]{x_i} \norm{x}^{2-n} \dd{x}. 
	\]
	Now, it is easily computed that $\pdv{x_i} \norm{x} = \frac{x_i}{\norm{x}}$, and therefore 
	\begin{align*}
	\pdv[2]{x_i} \norm{x}^{2-n} &= \pdv{x_i} (2-n) x_i \norm{x}^{-n} = (2-n) \qty(\norm{x}^{-n} - n x_i^2 \norm{x}^{-n-2}) \\
	&= (n-2)\norm{x}^{-n} \qty(n \qty(\frac{x_i}{\norm{x}})^2 - 1). 
	\end{align*}
\TODO finish
\end{solution}

\begin{question}
	Let $(f_k)$ be the sequence of smooth functions defined by $f_k(x) = \frac1\pi \frac k{(kx)^2 + 1}$. Prove that $f_k \to \delta_0$ in $\DD'(\RR)$. Compute the limits
	\begin{enumerate}[(a)]
		\item $\lim_{k\to\infty} k^2 x e^{-k^2 x^2}$,
		\item $\lim_{k\to\infty} k^3 e^{ikx}$,
		\item $\lim_{k\to\infty} \frac{\sin(kx)}{\pi x}$,
	\end{enumerate}
in $\DD'(\RR)$. 
\end{question}

\begin{solution}
	It is easily seen that every $f_k$ is locally integrable. Furthermore, noting that $\arctan(kx)' = \frac{k}{(kx)^2 + 1}$, we see by the dominated convergence theorem that
	\begin{align*}
		\lim_{k\to\infty}\ang{f_k, \phi} &= \frac1\pi \lim_{k\to\infty} \int_\RR \frac{k}{(kx)^2 + 1} \phi(x) \dd{x} = -\frac1\pi \lim_{k\to\infty} \int_\RR \arctan(kx) \phi'(x) \dd{x} \\
		&= -\frac1\pi \int_\RR \lim_{k\to\infty} \arctan(kx) \phi'(x) \dd{x} = -\frac1\pi \qty(2 \frac\pi2 \int_0^\infty \phi'(x) \dd{x} ) = \phi(0), 
	\end{align*}
which proves $f_k \to \delta_0$.

\begin{enumerate}[(a)]
	\item We have $- (\frac12 e^{-k^2x^2})' = k^2 x e^{-k^2 x^2}$ and therefore
	\begin{align*}
		\lim_{k\to\infty} \int_\RR k^2 x e^{-k^2x^2} \phi(x) \dd{x} &= \frac12 \lim_{k\to\infty}  \int_\RR  e^{-k^2 x^2} \phi'(x) \dd{x} = \frac12 \int_\RR \lim_{k\to\infty} e^{-k^2 x^2} \phi'(x) \dd{x} = 0,
	\end{align*}
so the sequence converges to 0 in $\DD'(\RR)$.

\item We have 
\begin{align*}
	\lim_{k\to\infty}  k^3 \int_\RR e^{ikx} \phi(x) \dd{x} = \lim_{k\to\infty} 2\pi k^3 \Cal F^{-1}[\phi](k). 
\end{align*}
Since $\phi$ is a Schwarz function, its inverse Fourier transform is also a Schwarz function, and in particular $k^3 \Cal F^{-1}[\phi](k) \to 0$ as $k \to \infty$.

\item This sequence converges to $\delta_0$, although I have no idea how to prove it.

\end{enumerate}
\end{solution}

\begin{question}
	Compute the limit
	\[
	\lim_{k\to\infty} \frac12 + \sum_{m=1}^k \cos(\pi m x)
	\]
	in $\DD'(-1, 1)$. 
\end{question}

\begin{solution}
	Note that \[
	f_k(x) \ceq 
	\frac12 + \sum_{m=1}^k \cos(\pi m x) = \frac12\qty(1 + e^{i \pi x} + e^{-i\pi x} + \dotsb + e^{i \pi m x} + e^{-i\pi m x}) = \frac12 \sum_{m=-k}^k (e^{i\pi x})^m.
	\]
	By viewing the last term as a geometric series $\frac12 e^{-ik\pi x} \sum_{m=0}^{2k+1} (e^{ix})^m$ we can compute that 
	\[
	f_k(x) = \frac{\sin((n + \frac12) \pi x)}{2\sin(\frac12\pi x)}. 
	\]
	It can be shown (???) that $f_k \to \delta_0$ in $\DD'(\RR)$. 
\end{solution}

\begin{question}
	We define the \emph{principal value} of $1/x$, written $\prv(1/x)$, by
	\[
	\ang*{\prv(1/x), \phi} = \lim_{\eps \to 0} \int_{\abs{x} > \eps} \frac{\phi(x)}{x} \dd{x} \quad\text{for all $\phi \in \DD(\RR)$}. 
	\]
	Prove that $\prv(1/x) \in \DD'(\RR)$ and that $\ord(\prv(1/x)) = 1$. Show that
	\[
	\lim_{\eps \to 0} \frac{1}{x - i\eps} = \prv\qty(\frac1x) + i\pi \delta_0 \quad\text{in $\DD'(\RR)$}. 
	\]
\end{question}

\begin{proof}
	First we must show that $\prv(1/x)$ is well-defined.  ??
\end{proof}

\begin{question}
	??
\end{question}
\begin{proof}
	??
\end{proof}

\begin{question}
	??
\end{question}

\begin{proof}
	??
\end{proof}

\begin{question}
	Define the distribution $u \in \EE'(\RR^3)$ by the locally integrable function $u(x) = \ind_{\abs{x}\leq 1}$. Prove that $-\sum_i x_i (\pdv{u}{x_i}) = \dd{\sigma_2}$ in $\EE'(\RR^3)$, where $\dd{\sigma_2}$ is the surface element on the sphere $S^2 \subseteq \RR^3$. 
\end{question}

\begin{proof}
	We have
	\begin{align*}
		\ang*{- \sum_i x_i \pdv{u}{x_i}, \phi} &= - \sum_i \ang*{\pdv{u}{x_i}, x_i\phi} = \sum_i \ang*{u, \pdv{x_i}(x_i \phi)} = \ang*{u, \sum_i \pdv{x_i}(x_i\phi)} \\
		&= \int_{\norm{x} \leq 1} \sum_i \pdv{x_i}(x_i\phi) \dd{x}
	\end{align*}
\end{proof}
\end{document}
